---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# panels

<!-- badges: start -->
<!-- badges: end -->

The goal of panels is to provide tools to model panel data.

## Installation

You can install the released version of panels from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("panels")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("mateuszwyszynski/panels")
```

## Basic Usage

```{r}
library(magrittr)
devtools::load_all()

set.seed(20)

# Prepare data
#
# Features are scaled and centralized around the mean.
# Then they are centralized around the mean within cross-sections
data_prepared <- panels::economic_growth[,1:7] %>%
  feature_standardization(timestamp_col = year, entity_col = country) %>%
  feature_standardization(timestamp_col = year, entity_col = country,
                          cross_sectional = TRUE, scale = FALSE)

# If needed track computation time
# begin<-Sys.time()

# Find optimal model space
#
# Parameters for each model are initialized with init_value. Then MLE for each
# model is searched numerically
model_space <-
  optimal_model_space(df = data_prepared, dep_var_col = gdp,
                      timestamp_col = year, entity_col = country,
                      init_value = 0.5, projection_matrix_const = TRUE)

# print(paste("Computation Time:", Sys.time()-begin))
# begin<-Sys.time() # Reset clock

# Compute intermediate BMA results
bma_result <- bma_summary(df = data_prepared, dep_var_col = gdp,
                          timestamp_col = year, entity_col = country,
                          model_space = model_space,
                          projection_matrix_const = TRUE)

# print(paste("Computation Time:", Sys.time()-begin))

# Summary for parameters of interest
regressors <- regressor_names(data_prepared, year, country, gdp)

bma_params_summary <- parameters_summary(
  regressors = regressors, bet = bma_result$bet, pvarh = bma_result$pvarh,
  pvarr = bma_result$pvarr, fy = bma_result$fy, fyt = bma_result$fyt,
  ppmsize = bma_result$ppmsize, cout = bma_result$cout, nts = bma_result$nts,
  pts = bma_result$pts, variables_n = bma_result$variables_n
  )
bma_params_summary
```

## Troubleshooting

1. Cannot install required packages / setup renv environment

Make sure to go through the displayed errors.
The problem might be connected to your OS environment.
E.g. you might see an information like the following:

```
Configuration failed to find one of freetype2 libpng libtiff-4 libjpeg. Try installing:
 * deb: libfreetype6-dev libpng-dev libtiff5-dev libjpeg-dev (Debian, Ubuntu, etc)
 * rpm: freetype-devel libpng-devel libtiff-devel libjpeg-devel (Fedora, CentOS, RHEL)
 * csw: libfreetype_dev libpng16_dev libtiff_dev libjpeg_dev (Solaris)
```

In such case you should first try installing the recommended packages.
With properly configured system environment everything should work fine.

## Comparison of methods with different projection matrix

The code snippets below show a problem which was present in the original
computations performed in the
[Growth Empirics in Panel Data Under Model Uncertainty and Weak Exogeneity](https://onlinelibrary.wiley.com/doi/abs/10.1002/jae.2429)
by Moral-Benito, i.e. an inconsistent use of the projection matrix.

More precisely, Moral-Benito used a simplified version of the likelihood,
derived in the paper
[Dynamic Panels with Predetermined Regressors: Likelihood-Based Estimation and Bayesian Averaging With an Application to Cross-Country Growth](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=1844186).
In the Appendix A.2 exact formulas for some of the parameters are obtained
in 3 steps through differentiation. At the end of the first step, just before
equation (23), a projection matrix is constructed from the matrix Z of initial
feature values. During the BMA analysis this matrix should be the same for all
considered models.

In the code obtained from Moral-Benito this was not the case.
If a linear relation between a feature and the dependent variable was set to
zero for a given model, a column relating to this feature inside the Z matrix
was also dropped. Most likely this was caused by the desire to optimize
computations. This column is not needed when one is multiplying the C and the Z
matrices, because some columns of the C matrix are automatically zero vectors.
However, the projection matrix introduced in the appendix is multiplied also by
other matrices whose columns do not have to be automatically zero (and most
likely are not).

This can be checked by comparing the value of the likelihood obtained with full
model and nested model implementations of the likelihood function. We will use
the MLE parameters obtained for one of the nested models as an example.
However, any parameters defining a nested model (i.e. with linear relation
parameters equal to zero) can be used as an example.

### MLE parameters example

Below, we extract the MLE parameters obtained for one of the nested models.
These parameters are then used to compute the value of the likelihood:
 - using two nested model implementations
 - using the full model implementation and setting respective `beta` and `phi_1`
 parameters to zero.

As an example nested model we use the model with `beta` and `phi_0` set to zero
for the `ish` feature. We compute three likelihoods:

- `no_ish_model_likelihood_1` where the likelihood is computed with the nested
model implementation of the likelihood function. Here the projection matrix
is computed as in the original Moral-Benito computations, i.e. with a column
corresponding to the `ish` feature dropped (`projection_matrix_const = FALSE`)
- `no_ish_model_likelihood_2` where the likelihood is computed with the nested
model implementation of the likelihood function. Here the projection matrix
is computed in a corrected manner, i.e. all columns are kept in the matrix
(`projection_matrix_const = TRUE`)
- `nested_no_ish_model_likelihood` where the likelihood is computed with the
full model implementation of the likelihood function. Here we use the same
values of the parameters, but we explicitly set the `beta` and `phi_0` to zero.
Hence, in this version some unnecessary multiplications by zero vectors are
computed.

```{r}
library(magrittr)
devtools::load_all()

set.seed(23)

data_prepared <- panels::economic_growth[,1:7] %>%
  feature_standardization(timestamp_col = year, entity_col = country) %>%
  feature_standardization(timestamp_col = year, entity_col = country,
                          cross_sectional = TRUE, scale = FALSE)

nested_version_of_no_ish_model_params <- economic_growth_ms[, 15]
no_ish_model_params <-
  nested_version_of_no_ish_model_params %>% stats::na.omit()
no_ish_model_likelihood_original <-
  SEM_likelihood(no_ish_model_params, data_prepared, year, country, gdp,
                 lin_related_regressors = c('sed', 'pgrw', 'pop'),
                 projection_matrix_const = FALSE)
no_ish_model_likelihood_corrected <-
  SEM_likelihood(no_ish_model_params, data_prepared, year, country, gdp,
                 lin_related_regressors = c('sed', 'pgrw', 'pop'),
                 projection_matrix_const = TRUE)

nested_version_of_no_ish_model_params['beta_ish'] = 0
nested_version_of_no_ish_model_params['phi_1_ish'] = 0

nested_no_ish_model_likelihood <-
  SEM_likelihood(nested_version_of_no_ish_model_params, data_prepared, year,
                 country, gdp,
                 lin_related_regressors = c('ish', 'sed', 'pgrw', 'pop'),
                 projection_matrix_const = TRUE)
```
One can clearly see that the result obtained with the corrected version matches
the result obtained with the full model implementation.

### Example with MLE parameters obtained with the incorrect implementation

To double check, we can use yet another example, this time using the parameters
obtained during MLE optimization performed with the incorrect implementation
of the likelihood function. We can see that the likelihood value obtained during
the optimization is inconsistent with the value obtained with the same
parameters used as an input for the full model.

```{r}
library(magrittr)
devtools::load_all()

set.seed(23)

data_prepared <- panels::economic_growth[,1:7] %>%
  feature_standardization(timestamp_col = year, entity_col = country) %>%
  feature_standardization(timestamp_col = year, entity_col = country,
                          cross_sectional = TRUE, scale = FALSE)

economic_growth_ms_2 <-
  optimal_model_space(df = data_prepared, dep_var_col = gdp,
                      timestamp_col = year, entity_col = country,
                      init_value = 0.5, projection_matrix_const = FALSE)

nested_version_of_no_ish_model_params <- economic_growth_ms_2[, 15]
no_ish_model_params <-
  nested_version_of_no_ish_model_params %>% stats::na.omit()
no_ish_model_likelihood_original <-
  SEM_likelihood(no_ish_model_params, data_prepared, year, country, gdp,
                 lin_related_regressors = c('sed', 'pgrw', 'pop'),
                 projection_matrix_const = FALSE)
no_ish_model_likelihood_corrected <-
  SEM_likelihood(no_ish_model_params, data_prepared, year, country, gdp,
                 lin_related_regressors = c('sed', 'pgrw', 'pop'),
                 projection_matrix_const = TRUE)

nested_version_of_no_ish_model_params['beta_ish'] = 0
nested_version_of_no_ish_model_params['phi_1_ish'] = 0

nested_no_ish_model_likelihood <-
  SEM_likelihood(nested_version_of_no_ish_model_params, data_prepared, year,
                 country, gdp,
                 lin_related_regressors = c('ish', 'sed', 'pgrw', 'pop'),
                 projection_matrix_const = FALSE)
```
Note that the value obtained with the corrected version matches the value
obtained with the nested model. This is because even in the original version,
in the full model a proper projection matrix is used, because all features are
present and so no columns are dropped in the Z matrix. As a result, from the
perspective of the corrected version we are just using some random parameters
to check if the value obtained with the nested model implementation and the full
model implementation is consistent. The only thing is that these parameters are
not MLEs, because they were obtained with an incorrect implementation.
